### 2018-05-24 mycv/cv/images.tex
tex files were added

\batchmode

\documentclass[11pt,a4paper]{awesome-cv}

\geometry{left=1.4cm, top=.8cm, right=1.4cm, bottom=1.8cm, footskip=.5cm}

\fontdir[fonts/]

\colorlet{awesome}{awesome-red}

\setbool{acvSectionColorHighlight}{true}

%
\renewcommand{\acvHeaderSocialSep}{\quad\textbar\quad}

\name{Youngung}{Jeong}
\position{Assistant professor}
\address{20 Changwondaehak-ro, Changwon, Gyeongnam, 51140, Republic of Korea}

\mobile{(+82) 10-9763-3022}
\email{yjeong@changwon.ac.kr}
\homepage{52-208}
\github{youngung}
\gitlab{youngung.jeong}

\usepackage[dvips]{color}

\pagecolor[gray]{.7}

\usepackage[]{inputenc}

\makeatletter

\makeatletter
\count@=\the\catcode`\_ \catcode`\_=8 
\newenvironment{tex2html_wrap}{}{}%
\catcode`\<=12\catcode`\_=\count@
\newcommand{\providedcommand}[1]{\expandafter\providecommand\csname #1\endcsname}%
\newcommand{\renewedcommand}[1]{\expandafter\providecommand\csname #1\endcsname{}%
  \expandafter\renewcommand\csname #1\endcsname}%
\newcommand{\newedenvironment}[1]{\newenvironment{#1}{}{}\renewenvironment{#1}}%
\let\newedcommand\renewedcommand
\let\renewedenvironment\newedenvironment
\makeatother
\let\mathon=$
\let\mathoff=$
\ifx\AtBeginDocument\undefined \newcommand{\AtBeginDocument}[1]{}\fi
\newbox\sizebox
\setlength{\hoffset}{0pt}\setlength{\voffset}{0pt}
\addtolength{\textheight}{\footskip}\setlength{\footskip}{0pt}
\addtolength{\textheight}{\topmargin}\setlength{\topmargin}{0pt}
\addtolength{\textheight}{\headheight}\setlength{\headheight}{0pt}
\addtolength{\textheight}{\headsep}\setlength{\headsep}{0pt}
\setlength{\textwidth}{349pt}
\newwrite\lthtmlwrite
\makeatletter
\let\realnormalsize=\normalsize
\global\topskip=2sp
\def\preveqno{}\let\real@float=\@float \let\realend@float=\end@float
\def\@float{\let\@savefreelist\@freelist\real@float}
\def\liih@math{\ifmmode$\else\bad@math\fi}
\def\end@float{\realend@float\global\let\@freelist\@savefreelist}
\let\real@dbflt=\@dbflt \let\end@dblfloat=\end@float
\let\@largefloatcheck=\relax
\let\if@boxedmulticols=\iftrue
\def\@dbflt{\let\@savefreelist\@freelist\real@dbflt}
\def\adjustnormalsize{\def\normalsize{\mathsurround=0pt \realnormalsize
 \parindent=0pt\abovedisplayskip=0pt\belowdisplayskip=0pt}%
 \def\phantompar{\csname par\endcsname}\normalsize}%
\def\lthtmltypeout#1{{\let\protect\string \immediate\write\lthtmlwrite{#1}}}%
\newcommand\lthtmlhboxmathA{\adjustnormalsize\setbox\sizebox=\hbox\bgroup\kern.05em }%
\newcommand\lthtmlhboxmathB{\adjustnormalsize\setbox\sizebox=\hbox to\hsize\bgroup\hfill }%
\newcommand\lthtmlvboxmathA{\adjustnormalsize\setbox\sizebox=\vbox\bgroup %
 \let\ifinner=\iffalse \let\)\liih@math }%
\newcommand\lthtmlboxmathZ{\@next\next\@currlist{}{\def\next{\voidb@x}}%
 \expandafter\box\next\egroup}%
\newcommand\lthtmlmathtype[1]{\gdef\lthtmlmathenv{#1}}%
\newcommand\lthtmllogmath{\dimen0\ht\sizebox \advance\dimen0\dp\sizebox
  \ifdim\dimen0>.95\vsize
   \lthtmltypeout{%
*** image for \lthtmlmathenv\space is too tall at \the\dimen0, reducing to .95 vsize ***}%
   \ht\sizebox.95\vsize \dp\sizebox\z@ \fi
  \lthtmltypeout{l2hSize %
:\lthtmlmathenv:\the\ht\sizebox::\the\dp\sizebox::\the\wd\sizebox.\preveqno}}%
\newcommand\lthtmlfigureA[1]{\let\@savefreelist\@freelist
       \lthtmlmathtype{#1}\lthtmlvboxmathA}%
\newcommand\lthtmlpictureA{\bgroup\catcode`\_=8 \lthtmlpictureB}%
\newcommand\lthtmlpictureB[1]{\lthtmlmathtype{#1}\egroup
       \let\@savefreelist\@freelist \lthtmlhboxmathB}%
\newcommand\lthtmlpictureZ[1]{\hfill\lthtmlfigureZ}%
\newcommand\lthtmlfigureZ{\lthtmlboxmathZ\lthtmllogmath\copy\sizebox
       \global\let\@freelist\@savefreelist}%
\newcommand\lthtmldisplayA{\bgroup\catcode`\_=8 \lthtmldisplayAi}%
\newcommand\lthtmldisplayAi[1]{\lthtmlmathtype{#1}\egroup\lthtmlvboxmathA}%
\newcommand\lthtmldisplayB[1]{\edef\preveqno{(\theequation)}%
  \lthtmldisplayA{#1}\let\@eqnnum\relax}%
\newcommand\lthtmldisplayZ{\lthtmlboxmathZ\lthtmllogmath\lthtmlsetmath}%
\newcommand\lthtmlinlinemathA{\bgroup\catcode`\_=8 \lthtmlinlinemathB}
\newcommand\lthtmlinlinemathB[1]{\lthtmlmathtype{#1}\egroup\lthtmlhboxmathA
  \vrule height1.5ex width0pt }%
\newcommand\lthtmlinlineA{\bgroup\catcode`\_=8 \lthtmlinlineB}%
\newcommand\lthtmlinlineB[1]{\lthtmlmathtype{#1}\egroup\lthtmlhboxmathA}%
\newcommand\lthtmlinlineZ{\egroup\expandafter\ifdim\dp\sizebox>0pt %
  \expandafter\centerinlinemath\fi\lthtmllogmath\lthtmlsetinline}
\newcommand\lthtmlinlinemathZ{\egroup\expandafter\ifdim\dp\sizebox>0pt %
  \expandafter\centerinlinemath\fi\lthtmllogmath\lthtmlsetmath}
\newcommand\lthtmlindisplaymathZ{\egroup %
  \centerinlinemath\lthtmllogmath\lthtmlsetmath}
\def\lthtmlsetinline{\hbox{\vrule width.1em \vtop{\vbox{%
  \kern.1em\copy\sizebox}\ifdim\dp\sizebox>0pt\kern.1em\else\kern.3pt\fi
  \ifdim\hsize>\wd\sizebox \hrule depth1pt\fi}}}
\def\lthtmlsetmath{\hbox{\vrule width.1em\kern-.05em\vtop{\vbox{%
  \kern.1em\kern0.8 pt\hbox{\hglue.17em\copy\sizebox\hglue0.8 pt}}\kern.3pt%
  \ifdim\dp\sizebox>0pt\kern.1em\fi \kern0.8 pt%
  \ifdim\hsize>\wd\sizebox \hrule depth1pt\fi}}}
\def\centerinlinemath{%
  \dimen1=\ifdim\ht\sizebox<\dp\sizebox \dp\sizebox\else\ht\sizebox\fi
  \advance\dimen1by.5pt \vrule width0pt height\dimen1 depth\dimen1 
 \dp\sizebox=\dimen1\ht\sizebox=\dimen1\relax}

\def\lthtmlcheckvsize{\ifdim\ht\sizebox<\vsize 
  \ifdim\wd\sizebox<\hsize\expandafter\hfill\fi \expandafter\vfill
  \else\expandafter\vss\fi}%
\providecommand{\selectlanguage}[1]{}%
\makeatletter \tracingstats = 1

\begin{document}
\pagestyle{empty}\thispagestyle{empty}\lthtmltypeout{}%
\lthtmltypeout{latex2htmlLength hsize=\the\hsize}\lthtmltypeout{}%
\lthtmltypeout{latex2htmlLength vsize=\the\vsize}\lthtmltypeout{}%
\lthtmltypeout{latex2htmlLength hoffset=\the\hoffset}\lthtmltypeout{}%
\lthtmltypeout{latex2htmlLength voffset=\the\voffset}\lthtmltypeout{}%
\lthtmltypeout{latex2htmlLength topmargin=\the\topmargin}\lthtmltypeout{}%
\lthtmltypeout{latex2htmlLength topskip=\the\topskip}\lthtmltypeout{}%
\lthtmltypeout{latex2htmlLength headheight=\the\headheight}\lthtmltypeout{}%
\lthtmltypeout{latex2htmlLength headsep=\the\headsep}\lthtmltypeout{}%
\lthtmltypeout{latex2htmlLength parskip=\the\parskip}\lthtmltypeout{}%
\lthtmltypeout{latex2htmlLength oddsidemargin=\the\oddsidemargin}\lthtmltypeout{}%
\makeatletter
\if@twoside\lthtmltypeout{latex2htmlLength evensidemargin=\the\evensidemargin}%
\else\lthtmltypeout{latex2htmlLength evensidemargin=\the\oddsidemargin}\fi%
\lthtmltypeout{}%
\makeatother
\setcounter{page}{1}
\onecolumn

% !!! IMAGES START HERE !!!

{\newpage\clearpage
\lthtmlfigureA{cventries49}%
\begin{cventries}
\par
\cventry
  {PhD} % Degree
  {POSTECH, Graduate Institute of Ferrous Technology} % Institution
  {Pohang, Republic of Korea} % Location
  {Mar. 2010 - Feb. 2014} % Date(s)
  {
  }
\par
\cventry
  {MS} % Degree
  {POSTECH, Graduate Institute of Ferrous Technology} % Institution
  {Pohang, Republic of Korea} % Location
  {Mar. 2008 - Feb. 2010} % Date(s)
  {
  }
\par
\cventry
  {BS} % Degree
  {Hanyang University, Materials Science and Engineering} % Institution
  {Seoul, Republic of Korea} % Location
  {Mar. 2001 - Feb. 2008} % Date(s)
  {
  }
  % {
  \par
\end{cventries}%
\lthtmlfigureZ
\lthtmlcheckvsize\clearpage}

{\newpage\clearpage
\lthtmlfigureA{cventries77}%
\begin{cventries}
  % ---------------------------------------------------------
  \cventry
  {Assistant professor} % Job title
  {Changwon National University}
  {Changwon, Republic of Korea} % Location
  {Mar. 2017, - present} % Date(s)
  {
  }
\par
\cventry
  {Post Doctorate Researcher} % Job title
  {POSTECH}
  {Pohang, Republic of Korea} % Location
  {Dec. 2016, - Feb. 2017} % Date(s)
  {
    \begin{cvitems} % Description(s) of tasks/responsibilities
    \item {Crystal plasticity modeling and simulations to predict bauschinger effects}
    \end{cvitems}
  }
\par
\cventry
  {Research Scientist} % Job title
  {Clemson University}
  {Greenville, SC, USA} % Location
  {Mar. 2016, - Nov. 2016} % Date(s)
  {
    \begin{cvitems} % Description(s) of tasks/responsibilities
    \item {Formability predictions using VPSC-FLD}
    \item {Abaqus UMAT development}
    \end{cvitems}
  }
\par
\cventry
  {Post Doctorate Researcher} % Job title
  {National Institute of Standards and Technology} % Organization
  {Gaithersburg, MD, USA} % Location
  {Jan. 2014, - Feb. 2016} % Date(s)
  {
    \begin{cvitems} % Description(s) of tasks/responsibilities
    \item {Conducted a series of experiments to obtain multiaxial constitutive behavior of an interstitial-free steel}
    \item {Measured multiaxial flow stress using X-ray diffraction for metal sheets subjected to various multiaxial loading conditions (DiffStress package)}
    \item {Performed the strain analysis using digital image correlation technique to determine the forming limit diagram of the IF steel}
    \item {Developed the VPSC-FLD model to predict forming limit diagram of engineering metal sheets (VPSC-FLD package)}
    \item {Developed VPSC-based model to link with continuum-scale phenomenological model (VPSC-RGVB-YLD forked from VPSC-FLD)}
    \end{cvitems}
    % \begin{cvsubentries}
    }
\par
\cventry
  {Research Affiliate} % Job title
  {Los Alamos National Laboratory} % Organization
  {Los Alamos, NM, USA} % Location
  {Apr. 2012, - Sep. 2012} % Date(s)
  {
    \begin{cvitems} % Description(s) of tasks/responsibilities
      {Leading role in implementing a phase transformation kinetics model into Elasto-ViscoPlastic Self-Consistent crystal plasticity model}
    \end{cvitems}
  }
\par
\cventry
  {Guest Researcher} % Job title
  {National Institute of Standards and Technology} % Organization
  {Gaithersburg, MD, USA} % Location
  {June. 2011, - Dec. 2011} % Date(s)
  {
    \begin{cvitems} % Description(s) of tasks/responsibilities
      {Conducted experiments to obtain multiaxial stress-strain measurements using digital image correlation and in-situ X-ray technique}
    \end{cvitems}
  }
\par
\end{cventries}%
\lthtmlfigureZ
\lthtmlcheckvsize\clearpage}

{\newpage\clearpage
\lthtmlfigureA{cvskills144}%
\begin{cvskills}
  \cvskill
  {Programming}
  {Python, Fortran, Bash script, C/C++, LaTeX, Matlab}
\par
\cvskill
  {Languages}
  {Korean, English}
\par
\cvskill
  {Experimental Mechanics}
  {Uniaxial tension, shear, hydraulic bulge test, biaxial tests using cruciform piece and Marciniak}
\par
\cvskill
  {Digital Image Correlation (DIC)}
  {VIC3D}
\par
\cvskill
  {Diffraction experiments}
  {Pole figure, crystallographic texture, phase fraction, residual stress measurements}
\par
\cvskill
  {Computer skills}
  {Linux, Git, Abaqus (UMAT and Python script), Parallel computation}
\par
\cvskill
  {Constutitive modelling}
  {Macro-mechanical description for anisotropic metals using anisotropic yield functions}
\par
\cvskill
  {Crystal plasticity}
  {Viscoplastic self-consistent (VPSC) and Elasto-viscoplastic self-consistent crystal plasticity models}
\par
\end{cvskills}%
\lthtmlfigureZ
\lthtmlcheckvsize\clearpage}

{\newpage\clearpage
\lthtmlfigureA{cventries181}%
\begin{cventries}
  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, F. Barlat, C. Tom\'{e}, W. Wen}
  {\href{http://dx.doi.org/10.1016/j.ijplas.2016.07.015}{A comparative study between micro- and macro-mechanical constitutive models developed for complex loading scenarios}}
  {International Journal of Plasticity}
  {Vol. 93 p212-228, 2017} % etcs
  {
  }
\par
\cventry
  {\underline{Y. Jeong}$^*$, T. Gn\"{a}upel-Herold, M. Iadicola, A. Creuziger} % Authors
  {\href{https://doi.org/10.1107/S1600576716013662}{Uncertainty in flow stress measurements using X-ray diffraction for sheet metals subjected to large plastic deformations}} % Title
  {Journal of Applied Crystallography} % Journals
  {Vol. 49 p1991-2004, 2016} % etcs
  {
  }
\par
\cventry
  {D. Steglich, \underline{Y. Jeong}$^*$}
  {\href{http://dx.doi.org/10.1016/j.ijmecsci.2016.08.013}{Texture-based forming limit prediction for Mg sheet alloys ZE10 and AZ31}}
  { International Journal of Mechanical Sciences} % Journals
  {Vol. 117 p102-114, 2016} % etcs
  {
  }
\par
\cventry
  {\underline{Y. Jeong}$^*$, M.-S. Pham, M. Iadicola, A. Creuziger, T. Foecke}
  {\href{http://dx.doi.org/10.1088/0965-0393/24/5/055005}{Forming limit prediction using a self-consistent crystal plasticity framework: a case study for BCC fiber textures}}
  {Modelling and Simulation in Materials Science and Engineering}
  {Vol. 24(5), 055002 (21 pp), 2016} % etcs
  {
  }
\par
\cventry
  {\underline{Y. Jeong}$^*$, M. Iadicola, T. Gn\"{a}upel-Herold, A. Creuziger}
  {\href{http://dx.doi.org/10.1016/j.actamat.2016.04.013}{Multiaxial constitutive behavior of an interstitial-free steel: measurements through X-ray and digital image correlation}}
  {Acta Materialia }
  {Vol. 112 p84-93, 2016}
  {
  }
\par
\cventry
  {H. Wang$^*$, \underline{Y. Jeong}, B. Clausen, Y. Liu, R. J. McCabe, F. Barlat, C. N. Tom\'{e}}
  {\href{http://dx.doi.org/10.1016/j.msea.2015.09.108}{Effect of martensitic phase transformation on the behavior of 304 austenitic stainless steel under tension}}
  {Materials Science and Engineering A }
  {Vol. 649 p174-183, 2016}
  {
  }
\par
\cventry
  {\underline{Y. Jeong}, T. Gn\"{a}upel-Herold, F. Barlat, M. Iadicola, A. Creuziger, M.-G. Lee$^*$}
  {\href{http://dx.doi.org/10.1016/j.ijplas.2014.06.009}{Evaluation of biaxial flow stress based on Elasto-Viscoplastic Self-Consistent analysis of X-ray Diffraction Measurements}}
  {International Journal of Plasticity}
  {Vol. 66 p103-118, 2015}
  {
  }
\par
\cventry
  {\underline{Y. Jeong}$^*$, F. Barlat, M.-G. Lee}
  {\href{http://dx.doi.org/10.1088/0965-0393/20/2/024009}{Application of crystal plasticity to an austenitic stainless steel}}
  {Modelling and Simulation in Materials Science and Engineering}
  {Vol. 20 p024009, 2012}
  {
  }
\par
\cventry
  {D. Steglich$^*$, \underline{Y. Jeong}, M. O. Andar, T. Kuwabara}
  {\href{http://dx.doi.org/10.1016/j.ijsolstr.2012.06.017}{Biaxial Deformation Behavior of AZ31 Magnesium Alloy: Crystal-Plasticity-Based Prediction and Experimental Validation}}
  {International Journal of Solids and Structure}
  {Vol. 49(25) p3551-3561, 2012}
  {
  }
\item
\par
\end{cventries}%
\lthtmlfigureZ
\lthtmlcheckvsize\clearpage}

{\newpage\clearpage
\lthtmlfigureA{cventries243}%
\begin{cventries}
  % ---------------------------------------------------------
\item
  \cventry
  {F. Barlat, \underline{Y. Jeong}, J. Ha, C Tom\'{e}, Myoung-Gyu Lee, W. Wen}
  {\href{https://dx.doi.org/10.4028/www.scientific.net/KEM.725.3}{Advances in Constitutive Modeling of Plasticity for Forming Applications}}
  {Key Engineering Materials}
  {Vol. 725, p3-14, 2017}
  {
  }
\par
\cventry
  {\underline{Y. Jeong}, F. Barlat, C. Tom\'{e}, W. Wen}
  {\href{http://dx.doi.org/10.1063/1.4963544}{Validation of Homogeneous Anisotropic Hardening Approach Based on Crystal Plasticity}}
  {AIP Conference Proceedings}
  {Vol. 1769, 160001, 2016}
  {
  }
\par
\cventry
  {\underline{Y. Jeong}, M.-S. Pham, M. Iadicola, A. Creuziger}
  {\href{https://doi.org/10.4028/www.scientific.net/KEM.651-653.193}{Forming limit predictions using a self-consistent crystal plasticity model: a parametric study}}
  {Key Engineering Materials}
  {Vol. 651 p193-198, 2015  }
  {
  }
\par
\cventry
  {\underline{Y. Jeong}, F. Barlat, M.-G. Lee}
  {\href{https://doi.org/10.4028/www.scientific.net/MSF.702-703.224}{Microstructural and crystallographic aspects of yield surface evolution}}
  {Materials Science Forum}
  {Vol. 702 p224-228, 2011}
  {
  }
\par
\cventry
  {\underline{Y. Jeong}, F. Barlat, M.-G. Lee}
  {\href{https://doi.org/10.4028/www.scientific.net/MSF.702-703.204}{Crystal Plasticity Predictions of Forward-Reverse Simple Shear Flow Stress}}
  {Materials Science Forum}
  {Vol. 702 p204-207, 2011}
  {
  }
\item
\par
\end{cventries}%
\lthtmlfigureZ
\lthtmlcheckvsize\clearpage}

\end{document}