### 2018-05-24 mycv/subs/fundings.tex

\cvsection{Fundings}
\begin{cvhonors}

  \cvhonor
  {Multi-scale crystal plasticity framework for multiphase metals under non-linear forming scenario} %Award
  {NRF-2017R1D1A1B03031052}%EVENT
  {National Science Foundation of Korea} %Location
  {2017-2020} %dates

\end{cvhonors}