### 2018-08-03 mycv/subs/fundings.tex
updates

\cvsection{Fundings}
\begin{cvhonors}

  \cvhonor
  {}
  {}
  {\href{http://www.motie.go.kr/www/main.do}{South Korean Ministry of Trade, Industry and Energy}}
  {2018-2021}

  \cvhonor
  {Multi-scale crystal plasticity framework for multiphase metals under non-linear forming scenario} %Award
  {NRF-2017R1D1A1B03031052}%EVENT
  {NSF of Korea} %Location
  {2017-2020} %dates

\end{cvhonors}