### 2019-08-15 mycv/subs/papers.tex
Updates

%-------------------------------------------------------------------------------
%	SECTION TITLE
%-------------------------------------------------------------------------------
\cvsection{Journal Articles}

%-------------------------------------------------------------------------------
%	CONTENT
%-------------------------------------------------------------------------------
\begin{cventries}
  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, C. N. Tom\'{e}}
  {Extension of the VPSC model to account for elasto-visco-plastic behavior using a perturbed viscoplastic approach}
  {Submitted}
  {}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {S. J. Park, \underline{Y. Jeong}, C. W. Kim, J. H. Lee, S. C. Cho, S. B. Lee, S. K. Lee, D. H. Kim, H. U. Hong$^*$}
  {\href{https://doi.org/10.1016/j.msea.2019.138260}{Superior tensile fracture strength of hot isostatically pressed TiC–steel metallic composite fabricated by a novel infiltration}}
  {Materials Science and Engineering: A}
  {Vol. 764(9), 2019}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, Timo Manninen}
  {\href{10.1007/s12540-019-00292-5}{Enhancement in viscoplastic self-consistent FLD prediction model and its application for austenitic and ferritic stainless steels}}
  {Metals and Materials International}
  {In Press}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {H. Kim, F Barlat, Y. Lee, S. Zaman, CS Lee, \underline{Y. Jeong}$^*$}
  {\href{https://doi.org/10.1016/j.ijplas.2018.07.010}{A crystal plasticity model for describing the anisotropic hardening behavior of steel sheets during strain-path changes}}
  {International Journal of Plasticity}
  {Vol. 111 p85-106, 2018}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, F. Barlat, C. Tom\'{e}, W. Wen}
  {\href{http://dx.doi.org/10.1016/j.ijplas.2016.07.015}{A comparative study between micro- and macro-mechanical constitutive models developed for complex loading scenarios}}
  {International Journal of Plasticity}
  {Vol. 93 p212-228, 2017} % etcs
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, T. Gn\"{a}upel-Herold, M. Iadicola, A. Creuziger} % Authors
  {\href{https://doi.org/10.1107/S1600576716013662}{Uncertainty in flow stress measurements using X-ray diffraction for sheet metals subjected to large plastic deformations}} % Title
  {Journal of Applied Crystallography} % Journals
  {Vol. 49 p1991-2004, 2016} % etcs
  {
  }

  % ---------------------------------------------------------
  \cventry
  {D. Steglich, \underline{Y. Jeong}$^*$}
  {\href{http://dx.doi.org/10.1016/j.ijmecsci.2016.08.013}{Texture-based forming limit prediction for Mg sheet alloys ZE10 and AZ31}}
  {International Journal of Mechanical Sciences} % Journals
  {Vol. 117 p102-114, 2016} % etcs
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, M.-S. Pham, M. Iadicola, A. Creuziger, T. Foecke}
  {\href{http://dx.doi.org/10.1088/0965-0393/24/5/055005}{Forming limit prediction using a self-consistent crystal plasticity framework: a case study for BCC fiber textures}}
  {Modelling and Simulation in Materials Science and Engineering}
  {Vol. 24(5), 055002 (21 pp), 2016} % etcs
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, M. Iadicola, T. Gn\"{a}upel-Herold, A. Creuziger}
  {\href{http://dx.doi.org/10.1016/j.actamat.2016.04.013}{Multiaxial constitutive behavior of an interstitial-free steel: measurements through X-ray and digital image correlation}}
  {Acta Materialia }
  {Vol. 112 p84-93, 2016}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {H. Wang$^*$, \underline{Y. Jeong}, B. Clausen, Y. Liu, R. J. McCabe, F. Barlat, C. N. Tom\'{e}}
  {\href{http://dx.doi.org/10.1016/j.msea.2015.09.108}{Effect of martensitic phase transformation on the behavior of 304 austenitic stainless steel under tension}}
  {Materials Science and Engineering A }
  {Vol. 649 p174-183, 2016}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}, T. Gn\"{a}upel-Herold, F. Barlat, M. Iadicola, A. Creuziger, M.-G. Lee$^*$}
  {\href{http://dx.doi.org/10.1016/j.ijplas.2014.06.009}{Evaluation of biaxial flow stress based on Elasto-Viscoplastic Self-Consistent analysis of X-ray Diffraction Measurements}}
  {International Journal of Plasticity}
  {Vol. 66 p103-118, 2015}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, F. Barlat, M.-G. Lee}
  {\href{http://dx.doi.org/10.1088/0965-0393/20/2/024009}{Application of crystal plasticity to an austenitic stainless steel}}
  {Modelling and Simulation in Materials Science and Engineering}
  {Vol. 20 p024009, 2012}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {D. Steglich$^*$, \underline{Y. Jeong}, M. O. Andar, T. Kuwabara}
  {\href{http://dx.doi.org/10.1016/j.ijsolstr.2012.06.017}{Biaxial Deformation Behavior of AZ31 Magnesium Alloy: Crystal-Plasticity-Based Prediction and Experimental Validation}}
  {International Journal of Solids and Structure}
  {Vol. 49(25) p3551-3561, 2012}
  {
  }
\item

\end{cventries}

%-------------------------------------------------------------------------------
%	SECTION TITLE
%-------------------------------------------------------------------------------
\cvsection{Conference proceedings}

%-------------------------------------------------------------------------------
%	CONTENT
%-------------------------------------------------------------------------------
\begin{cventries}
  % ---------------------------------------------------------
\item
    \cventry
  {\underline{Y. Jeong}$^*$, Timo Manninen}
  {\href{http://iopscience.iop.org/article/10.1088/1742-6596/1063/1/012020}{Formability predictions and measurement of 316L stainless steel using self-consistent crystal plasticity}}
  {Journal of Physics: Conference Series}
  {Vol. 150673, 2018}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, S. Panich}
  {\href{https://doi.org/10.1016/j.promfg.2018.07.209}{Forming limits of dual phase steels using crystal plasticity in conjunction with MK approach}}
  {Procedia Manufacturing}
  {Vol. 15, 2018}
  {
  }

  \cventry
  {D. Steglich \underline{Y. Jeong}}
  {\href{https://doi.org/10.1063/1.5007958}{Texture-based formability prediction for Mg wrought alloys ZE10 and AZ31}}
  {AIP Conference Proceedings}
  {Vol. 1896, 020001, 2017}
  {
  }

  \cventry
  {F. Barlat, \underline{Y. Jeong}, J. Ha, C Tom\'{e}, Myoung-Gyu Lee, W. Wen}
  {\href{https://dx.doi.org/10.4028/www.scientific.net/KEM.725.3}{Advances in Constitutive Modeling of Plasticity for Forming Applications}}
  {Key Engineering Materials}
  {Vol. 725, p3-14, 2017}
  {
  }

  \cventry
  {\underline{Y. Jeong}, F. Barlat, C. Tom\'{e}, W. Wen}
  {\href{http://dx.doi.org/10.1063/1.4963544}{Validation of Homogeneous Anisotropic Hardening Approach Based on Crystal Plasticity}}
  {AIP Conference Proceedings}
  {Vol. 1769, 160001, 2016}
  {
  }

  \cventry
  {\underline{Y. Jeong}, M.-S. Pham, M. Iadicola, A. Creuziger}
  {\href{https://doi.org/10.4028/www.scientific.net/KEM.651-653.193}{Forming limit predictions using a self-consistent crystal plasticity model: a parametric study}}
  {Key Engineering Materials}
  {Vol. 651 p193-198, 2015  }
  {
  }

  \cventry
  {\underline{Y. Jeong}, F. Barlat, M.-G. Lee}
  {\href{https://doi.org/10.4028/www.scientific.net/MSF.702-703.224}{Microstructural and crystallographic aspects of yield surface evolution}}
  {Materials Science Forum}
  {Vol. 702 p224-228, 2011}
  {
  }

  \cventry
  {\underline{Y. Jeong}, F. Barlat, M.-G. Lee}
  {\href{https://doi.org/10.4028/www.scientific.net/MSF.702-703.204}{Crystal Plasticity Predictions of Forward-Reverse Simple Shear Flow Stress}}
  {Materials Science Forum}
  {Vol. 702 p204-207, 2011}
  {
  }
\item

\end{cventries}