### 2020-06-29 mycv/subs/skills.tex
Updates in diffraction.ipynb

%-------------------------------------------------------------------------------
%	SECTION TITLE
%-------------------------------------------------------------------------------
\cvsection{Skills}

%-------------------------------------------------------------------------------
%	CONTENT
%-------------------------------------------------------------------------------
\begin{cvskills}
  \cvskill
  {Programming}
  {Python, Fortran, Bash script, C/C++, LaTeX, Matlab}

  \cvskill
  {Languages}
  {Korean, English}

  \cvskill
  {Experimental Mechanics}
  {Uniaxial tension, shear, hydraulic bulge test, biaxial tests using cruciform piece and Marciniak}

  \cvskill
  {Digital Image Correlation (DIC)}
  {VIC3D, DICE}

  \cvskill
  {Diffraction experiments}
  {Pole figure, crystallographic texture, phase fraction, residual stress measurements}

  \cvskill
  {Computer skills}
  {Linux, Git, Parallel computation}

  \cvskill
  {Constitutive modelling}
  {Macro-mechanical description on anisotropic metals using anisotropic yield functions}

  \cvskill
  {Crystal plasticity}
  {Visco-plastic / Elasto-visco-plastic self-consistent crystal plasticity models}

\end{cvskills}