### 2023-10-10 mycv/subs/papers.tex
updates in papers

%-------------------------------------------------------------------------------
%	SECTION TITLE
%-------------------------------------------------------------------------------
\cvsection{Journal Articles}

%-------------------------------------------------------------------------------
%	CONTENT
%-------------------------------------------------------------------------------
\begin{cventries}
  % ---------------------------------------------------------

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, C. N. Tom\'{e}}
  {Thermal ratcheting of uranium simulated by a thermo-elasto-visco-plastic self-consistent polycrystal model}
  {submitted}
  {}
  {
  }
  % ---------------------------------------------------------
  \cventry
  {B. Jeon, S.-Y. Lee, J. Lee, \underline{Y. Jeong}$^*$}
  {Direct application of elasto-viscoplastic self-consistent crystal plasticity model to U-draw bending and springback of dual-phase high strength steel}
  {in preparation}
  {}
  {
  }
  % ---------------------------------------------------------
  \cventry
  {B. Jeon, M.-S. Lee, T.-S. Jun, \underline{Y. Jeong}$^*$}
  {Temperature-dependent behavior of CP-Ti interpreted via self-consistent crystal plasticity simulation}
  {submitted}
  {}
  {
  }
  % ---------------------------------------------------------
  \cventry
  {B. Jeon, M.-S. Kim, S. Choi, \underline{Y. Jeong}$^*$}
  {\href{https://doi.org/10.1016/j.jma.2022.10.017}{Finite element analysis using elasto-visco-plastic self-consistent polycrystal model for E-form Mg sheet subjected to bending}}
  {Journal of Magnesium and Alloys}
  {Vol. 11(4), p1393-1407, 2023}
  {
  }
  % ---------------------------------------------------------
  \cventry
  {M. Joo, M.-S. Wi, S.-Y. Yoon, S.-Y. Lee, F. Barlat, C N. Tom\'{e}, B. Jeon, \underline{Y. Jeong}$^*$}
  {\href{https://doi.org/10.1016/j.ijmecsci.2022.107796}{A crystal plasticity finite element analysis on the effect of prestrain on springback}}
  {International Journal of Mechanical Sciences}
  {Vol. 237, 107796, 2023}
  {
  }
  % ---------------------------------------------------------
  \cventry
  {Lalit Kauschik, Ki-Seong Park, Jeong-Gyun Kim, Jae-Seong Lee, \underline{Youngung Jeong}, Shi-Hoon Choi$^*$}
  {\href{https://doi.org/10.1016/j.ijplas.2022.103467}{Reconstructing orientation data from the images of IPF maps and ODF sections extracted from the literature: A data-collection method for machine learning}}
  {International Journal of Plasticity}
  {Vol. 159, 103467, 2022}
  {
  }
  % ---------------------------------------------------------
  \cventry
  {J. Park, \underline{Y. Jeong}$^*$}
  {\href{http://dx.doi.org/10.3365/KJMM.2022.60.8.607}{Prediction and validation of stress triaxiality assisted by elasto-visco-plastic polycrystal model}}
  {Korean Journal of Metals and Materials}
  {Vol. 60 (8), 607-618, 2022}
  {
  }
  % ---------------------------------------------------------
  \cventry
  {M.S. Lee, T. Kawasaki, T. Yamashita, S. Harjo, Y.T. Hyun, \underline{Y. Jeong}, T. S. Jun$^*$}
  {\href{https://doi.org/10.1038/s41598-022-07640-3}{In-situ neutron diffraction study of lattice deformation behaviour of commercially pure titanium at cryogenic temperature}}
  {Scientific Reports}
  {Vol. 12 (1), 1-10, 2022}
  {
  }
  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, B. Jeon, C N. Tom\'{e}}
  {\href{https://doi.org/10.1016/j.ijplas.2021.103110}{Finite element analysis using an incremental elasto-visco-plastic self-consistent polycrystal model: FE simulations on Zr and low-carbon steel subjected to bending, stress-relaxation, and unloading}}
  {International Journal of Plasticity}
  {Vol. 147, 103110, 2021}
  {
  }
  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, C. N. Tom\'{e}}
  {\href{https://doi.org/10.1016/j.ijplas.2020.102812}{An efficient elasto-visco-plastic self-consistent formulation: Application to steel subjected to loading path changes}}
  {International Journal of Plasticity}
  {Vol. 135, 102812, 2020}
  {
  }
  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, D. Steglich}
  {\href{https://doi.org/10.1016/j.ijmecsci.2020.105680}{Modelling-assisted description of anisotropic edge failure in magnesium sheet alloy under mixed-mode loading}}
  {International Journal of Mechanical Sciences}
  {Vol. 181, 105680, 2020}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, C. N. Tom\'{e}}
  {\href{https://doi.org/10.1088/1361-651X/ab4b66}{Extension of the VPSC model to account for elasto-visco-plastic behavior using a perturbed viscoplastic approach}}
  {Modelling and Simulation in Materials Science and Engineering}
  {Vol. 27(8) 085013, 2019}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {S. J. Park, \underline{Y. Jeong}, C. W. Kim, J. H. Lee, S. C. Cho, S. B. Lee, S. K. Lee, D. H. Kim, H. U. Hong$^*$}
  {\href{https://doi.org/10.1016/j.msea.2019.138260}{Superior tensile fracture strength of hot isostatically pressed TiC–steel metallic composite fabricated by a novel infiltration}}
  {Materials Science and Engineering: A}
  {Vol. 764(9), 2019}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, Timo Manninen}
  {\href{https://doi.org/10.1007/s12540-019-00292-5}{Enhancement in viscoplastic self-consistent FLD prediction model and its application for austenitic and ferritic stainless steels}}
  {Metals and Materials International}
  {Vol. 25(6) pp1548-1563, 2019}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {H. Kim, F Barlat, Y. Lee, S. Zaman, CS Lee, \underline{Y. Jeong}$^*$}
  {\href{https://doi.org/10.1016/j.ijplas.2018.07.010}{A crystal plasticity model for describing the anisotropic hardening behavior of steel sheets during strain-path changes}}
  {International Journal of Plasticity}
  {Vol. 111 p85-106, 2018}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, F. Barlat, C. N. Tom\'{e}, W. Wen}
  {\href{http://dx.doi.org/10.1016/j.ijplas.2016.07.015}{A comparative study between micro- and macro-mechanical constitutive models developed for complex loading scenarios}}
  {International Journal of Plasticity}
  {Vol. 93 p212-228, 2017} % etcs
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, T. Gn\"{a}upel-Herold, M. Iadicola, A. Creuziger} % Authors
  {\href{https://doi.org/10.1107/S1600576716013662}{Uncertainty in flow stress measurements using X-ray diffraction for sheet metals subjected to large plastic deformations}} % Title
  {Journal of Applied Crystallography} % Journals
  {Vol. 49 p1991-2004, 2016} % etcs
  {
  }

  % ---------------------------------------------------------
  \cventry
  {D. Steglich, \underline{Y. Jeong}$^*$}
  {\href{http://dx.doi.org/10.1016/j.ijmecsci.2016.08.013}{Texture-based forming limit prediction for Mg sheet alloys ZE10 and AZ31}}
  {International Journal of Mechanical Sciences} % Journals
  {Vol. 117 p102-114, 2016} % etcs
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, M.-S. Pham, M. Iadicola, A. Creuziger, T. Foecke}
  {\href{http://dx.doi.org/10.1088/0965-0393/24/5/055005}{Forming limit prediction using a self-consistent crystal plasticity framework: a case study for BCC fiber textures}}
  {Modelling and Simulation in Materials Science and Engineering}
  {Vol. 24(5), 055002 (21 pp), 2016} % etcs
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, M. Iadicola, T. Gn\"{a}upel-Herold, A. Creuziger}
  {\href{http://dx.doi.org/10.1016/j.actamat.2016.04.013}{Multiaxial constitutive behavior of an interstitial-free steel: measurements through X-ray and digital image correlation}}
  {Acta Materialia }
  {Vol. 112 p84-93, 2016}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {H. Wang$^*$, \underline{Y. Jeong}, B. Clausen, Y. Liu, R. J. McCabe, F. Barlat, C. N. Tom\'{e}}
  {\href{http://dx.doi.org/10.1016/j.msea.2015.09.108}{Effect of martensitic phase transformation on the behavior of 304 austenitic stainless steel under tension}}
  {Materials Science and Engineering A }
  {Vol. 649 p174-183, 2016}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}, T. Gn\"{a}upel-Herold, F. Barlat, M. Iadicola, A. Creuziger, M.-G. Lee$^*$}
  {\href{http://dx.doi.org/10.1016/j.ijplas.2014.06.009}{Evaluation of biaxial flow stress based on Elasto-Viscoplastic Self-Consistent analysis of X-ray Diffraction Measurements}}
  {International Journal of Plasticity}
  {Vol. 66 p103-118, 2015}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, F. Barlat, M.-G. Lee}
  {\href{http://dx.doi.org/10.1088/0965-0393/20/2/024009}{Application of crystal plasticity to an austenitic stainless steel}}
  {Modelling and Simulation in Materials Science and Engineering}
  {Vol. 20 p024009, 2012}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {D. Steglich$^*$, \underline{Y. Jeong}, M. O. Andar, T. Kuwabara}
  {\href{http://dx.doi.org/10.1016/j.ijsolstr.2012.06.017}{Biaxial Deformation Behavior of AZ31 Magnesium Alloy: Crystal-Plasticity-Based Prediction and Experimental Validation}}
  {International Journal of Solids and Structure}
  {Vol. 49(25) p3551-3561, 2012}
  {
  }
\item

\end{cventries}

%-------------------------------------------------------------------------------
%	SECTION TITLE
%-------------------------------------------------------------------------------
\cvsection{Conference proceedings}

%-------------------------------------------------------------------------------
%	CONTENT
%-------------------------------------------------------------------------------
\begin{cventries}

\item
  % ---------------------------------------------------------
  \cventry
  {Carlos N. Tome$^*$, \underline{Y. Jeong}}
  {\href{https://doi.org/10.1007/978-3-031-40920-2_68}{Thermal Ratcheting of Uranium Simulated with a Thermo-Elasto-Visco-Plastic Polycrystal Model}}
  {Proceedings of the 14th International Conference on the Technology of Plasticity - Current Trends in the Technology of Plasticity}
  {}
  {
  }
  % ---------------------------------------------------------
  \cventry
  {B. Jeon, \underline{Y. Jeong}$^*$}
  {\href{https://doi.org/10.1007/978-3-031-40920-2_69}{Interpretation of the Unloading Non-linearity in Dual-Phase 980 Steel Using an Elasto-Visco-Plastic Self-consistent Polycrystal Model}}
  {Proceedings of the 14th International Conference on the Technology of Plasticity - Current Trends in the Technology of Plasticity}
  {}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, Timo Manninen}
  {\href{http://iopscience.iop.org/article/10.1088/1742-6596/1063/1/012020}{Formability predictions and measurement of 316L stainless steel using self-consistent crystal plasticity}}
  {Journal of Physics: Conference Series}
  {Vol. 150673, 2018}
  {
  }

  % ---------------------------------------------------------
  \cventry
  {\underline{Y. Jeong}$^*$, S. Panich}
  {\href{https://doi.org/10.1016/j.promfg.2018.07.209}{Forming limits of dual phase steels using crystal plasticity in conjunction with MK approach}}
  {Procedia Manufacturing}
  {Vol. 15, 2018}
  {
  }

  \cventry
  {D. Steglich \underline{Y. Jeong}}
  {\href{https://doi.org/10.1063/1.5007958}{Texture-based formability prediction for Mg wrought alloys ZE10 and AZ31}}
  {AIP Conference Proceedings}
  {Vol. 1896, 020001, 2017}
  {
  }

  \cventry
  {F. Barlat, \underline{Y. Jeong}, J. Ha, C. N. Tom\'{e}, M.-G. Lee, W. Wen}
  {\href{https://dx.doi.org/10.4028/www.scientific.net/KEM.725.3}{Advances in Constitutive Modeling of Plasticity for Forming Applications}}
  {Key Engineering Materials}
  {Vol. 725, p3-14, 2017}
  {
  }

  \cventry
  {\underline{Y. Jeong}, F. Barlat, C. N. Tom\'{e}, W. Wen}
  {\href{http://dx.doi.org/10.1063/1.4963544}{Validation of Homogeneous Anisotropic Hardening Approach Based on Crystal Plasticity}}
  {AIP Conference Proceedings}
  {Vol. 1769, 160001, 2016}
  {
  }

  \cventry
  {\underline{Y. Jeong}, M.-S. Pham, M. Iadicola, A. Creuziger}
  {\href{https://doi.org/10.4028/www.scientific.net/KEM.651-653.193}{Forming limit predictions using a self-consistent crystal plasticity model: a parametric study}}
  {Key Engineering Materials}
  {Vol. 651 p193-198, 2015  }
  {
  }

  \cventry
  {\underline{Y. Jeong}, F. Barlat, M.-G. Lee}
  {\href{https://doi.org/10.4028/www.scientific.net/MSF.702-703.224}{Microstructural and crystallographic aspects of yield surface evolution}}
  {Materials Science Forum}
  {Vol. 702 p224-228, 2011}
  {
  }

  \cventry
  {\underline{Y. Jeong}, F. Barlat, M.-G. Lee}
  {\href{https://doi.org/10.4028/www.scientific.net/MSF.702-703.204}{Crystal Plasticity Predictions of Forward-Reverse Simple Shear Flow Stress}}
  {Materials Science Forum}
  {Vol. 702 p204-207, 2011}
  {
  }
\item

\end{cventries}