### 2025-03-16 mycv/mylatex_cv.tex
snapshot of latex cv

\documentclass[10pt, letterpaper]{article}

% Packages:
\usepackage[
    ignoreheadfoot, % set margins without considering header and footer
    top=2 cm, % seperation between body and page edge from the top
    bottom=2 cm, % seperation between body and page edge from the bottom
    left=2 cm, % seperation between body and page edge from the left
    right=2 cm, % seperation between body and page edge from the right
    footskip=1.0 cm, % seperation between body and footer
    % showframe % for debugging
]{geometry} % for adjusting page geometry
\usepackage{titlesec} % for customizing section titles
\usepackage{tabularx} % for making tables with fixed width columns
\usepackage{array} % tabularx requires this
\usepackage[dvipsnames]{xcolor} % for coloring text
\definecolor{primaryColor}{RGB}{0, 0, 0} % define primary color
\usepackage{enumitem} % for customizing lists
\usepackage{fontawesome5} % for using icons
\usepackage{amsmath} % for math
\usepackage[
    pdftitle={John Doe's CV},
    pdfauthor={John Doe},
    pdfcreator={LaTeX with RenderCV},
    colorlinks=true,
    urlcolor=primaryColor
]{hyperref} % for links, metadata and bookmarks
\usepackage[pscoord]{eso-pic} % for floating text on the page
\usepackage{calc} % for calculating lengths
\usepackage{bookmark} % for bookmarks
\usepackage{lastpage} % for getting the total number of pages
\usepackage{changepage} % for one column entries (adjustwidth environment)
\usepackage{paracol} % for two and three column entries
\usepackage{ifthen} % for conditional statements
\usepackage{needspace} % for avoiding page brake right after the section title
\usepackage{iftex} % check if engine is pdflatex, xetex or luatex

% Ensure that generate pdf is machine readable/ATS parsable:
\ifPDFTeX
    \input{glyphtounicode}
    \pdfgentounicode=1
    \usepackage[T1]{fontenc}
    \usepackage[utf8]{inputenc}
    \usepackage{lmodern}
\fi

\usepackage{charter}

% Some settings:
\raggedright
\AtBeginEnvironment{adjustwidth}{\partopsep0pt} % remove space before adjustwidth environment
\pagestyle{empty} % no header or footer
\setcounter{secnumdepth}{0} % no section numbering
\setlength{\parindent}{0pt} % no indentation
\setlength{\topskip}{0pt} % no top skip
\setlength{\columnsep}{0.15cm} % set column seperation
\pagenumbering{gobble} % no page numbering

\titleformat{\section}{\needspace{4\baselineskip}\bfseries\large}{}{0pt}{}[\vspace{1pt}\titlerule]

\titlespacing{\section}{
    % left space:
    -1pt
}{
    % top space:
    0.3 cm
}{
    % bottom space:
    0.2 cm
} % section title spacing

\renewcommand\labelitemi{$\vcenter{\hbox{\small$\bullet$}}$} % custom bullet points
\newenvironment{highlights}{
    \begin{itemize}[
        topsep=0.10 cm,
        parsep=0.10 cm,
        partopsep=0pt,
        itemsep=0pt,
        leftmargin=0 cm + 10pt
    ]
}{
    \end{itemize}
} % new environment for highlights

\newenvironment{highlightsforbulletentries}{
    \begin{itemize}[
        topsep=0.10 cm,
        parsep=0.10 cm,
        partopsep=0pt,
        itemsep=0pt,
        leftmargin=10pt
    ]
}{
    \end{itemize}
} % new environment for highlights for bullet entries

\newenvironment{onecolentry}{
    \begin{adjustwidth}{
        0 cm + 0.00001 cm
    }{
        0 cm + 0.00001 cm
    }
}{
    \end{adjustwidth}
} % new environment for one column entries

\newenvironment{twocolentry}[2][]{
    \onecolentry
    \def\secondColumn{#2}
    \setcolumnwidth{\fill, 4.5 cm}
    \begin{paracol}{2}
}{
    \switchcolumn \raggedleft \secondColumn
    \end{paracol}
    \endonecolentry
} % new environment for two column entries

\newenvironment{threecolentry}[3][]{
    \onecolentry
    \def\thirdColumn{#3}
    \setcolumnwidth{, \fill, 4.5 cm}
    \begin{paracol}{3}
    {\raggedright #2} \switchcolumn
}{
    \switchcolumn \raggedleft \thirdColumn
    \end{paracol}
    \endonecolentry
} % new environment for three column entries

\newenvironment{header}{
    \setlength{\topsep}{0pt}\par\kern\topsep\centering\linespread{1.5}
}{
    \par\kern\topsep
} % new environment for the header

\newcommand{\placelastupdatedtext}{% \placetextbox{<horizontal pos>}{<vertical pos>}{<stuff>}
  \AddToShipoutPictureFG*{% Add <stuff> to current page foreground
    \put(
        \LenToUnit{\paperwidth-2 cm-0 cm+0.05cm},
        \LenToUnit{\paperheight-1.0 cm}
    ){\vtop{{\null}\makebox[0pt][c]{
        \small\color{gray}\textit{Last updated in September 2024}\hspace{\widthof{Last updated in September 2024}}
    }}}%
  }%
}%

% save the original href command in a new command:
\let\hrefWithoutArrow\href

% new command for external links:

\begin{document}
    \newcommand{\AND}{\unskip
        \cleaders\copy\ANDbox\hskip\wd\ANDbox
        \ignorespaces
    }
    \newsavebox\ANDbox
    \sbox\ANDbox{$|$}

    \begin{header}
        \fontsize{25 pt}{25 pt}\selectfont Youngung Jeong

        \vspace{5 pt}

        \normalsize
        \mbox{School of Materials Science and Engineering, Changwon National University}%
        \kern 5.0 pt%
        \AND%
        \kern 5.0 pt%
        \mbox{\hrefWithoutArrow{mailto:yjeong@changwon.ac.kr}{yjeong@changwon.ac.kr}}%
        \kern 5.0 pt%
        \AND%
        \kern 5.0 pt%
        \mbox{\hrefWithoutArrow{tel:+82-55-214-3694}{+82-55-214-3694}}%
        \kern 5.0 pt%
        \AND%
        \kern 5.0 pt%
        \mbox{\hrefWithoutArrow{https://youngung.github.io/}{youngung.github.io}}%
        \kern 5.0 pt%
        \AND%
        \kern 5.0 pt%
        \mbox{\hrefWithoutArrow{https://github.com/youngung}{github.com/youngung}}%
    \end{header}

    \vspace{5 pt - 0.3 cm}

%    \section{Welcome to RenderCV!}
%        \begin{onecolentry}
%            \href{https://rendercv.com}{RenderCV} is a LaTeX-based CV/resume version-control and maintenance app. It allows you to create a high-quality CV or resume as a PDF file from a YAML file, with \textbf{Markdown syntax support} and \textbf{complete control over the LaTeX code}.
%        \end{onecolentry}
%        \vspace{0.2 cm}
%        \begin{onecolentry}
%            The boilerplate content was inspired by \href{https://github.com/dnl-blkv/mcdowell-cv}{Gayle McDowell}.
%        \end{onecolentry}
%    \section{Quick Guide}
%    \begin{onecolentry}
%        \begin{highlightsforbulletentries}
%        \item Each section title is arbitrary and each section contains a list of entries.
%        \item There are 7 unique entry types: \textit{BulletEntry}, \textit{TextEntry}, \textit{EducationEntry}, \textit{ExperienceEntry}, \textit{NormalEntry}, \textit{PublicationEntry}, and \textit{OneLineEntry}.
%        \item Select a section title, pick an entry type, and start writing your section!
%        \item \href{https://docs.rendercv.com/user_guide/}{Here}, you can find a comprehensive user guide for RenderCV.
%        \end{highlightsforbulletentries}
%    \end{onecolentry}

    \section{Education}
        \begin{twocolentry}{Mar 2001 – Feb 2008}
            \textbf{Hanyang University}, BS, Department of Materials Science and Engineering
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{Mar 2008 – Feb 2010}
            \textbf{POSTECH}, MS, Graduate Institute of Ferrous Technology (supervisor: F. Barlat)
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{Mar 2010 – Feb 2014}
            \textbf{POSTECH}, PhD, Graduate Institute of Ferrous Technology (supervisor: F. Barlat)
        \end{twocolentry}
        \vspace{0.10 cm}

    \section{Experience}
        \begin{twocolentry}{Mar 2017 – Present}
            \textbf{Assistant, associate professor}, Changwon National University, ROK
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{July 2024 – Aug 2024}
            \textbf{Short term visitor}, Los Alamos National Laboratory, NM, USA
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{Feb 2022 – Feb 2024}
            \textbf{Guest Scientist (offsite)}, Los Alamos National Laboratory, NM, USA
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{Dec 2016 – Feb 2017}
            \textbf{Postdoc}, POSTECH, ROK
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{Mar 2016 – Nov 2016}
            \textbf{Research Scientist}, Clemson university, SC, USA
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{Mar 2014 – Feb 2016}
            \textbf{Postdoc}, NIST, MD, USA
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{Apr 2012 – Sep 2012}
            \textbf{Research Affiliate}, Los Alamos National Laboratory, NM, USA
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{June 2011 – Dec 2011}
            \textbf{Guest Researcher}, NIST, MD, USA
        \end{twocolentry}
        \vspace{0.10 cm}

%        \begin{twocolentry}{
%            June 2003 – Aug 2003
%        }
%            \textbf{Software Engineer Intern}, Microsoft -- Redmond, WA\end{twocolentry}

%        \vspace{0.10 cm}
%        \begin{onecolentry}
%            \begin{highlights}
%                \item Designed a UI for the VS open file switcher (Ctrl-Tab) and extended it to tool windows
%                \item Created a service to provide gradient across VS and VS add-ins, optimizing its performance via caching
%                \item Built an app to compute the similarity of all methods in a codebase, reducing the time from $\mathcal{O}(n^2)$ to $\mathcal{O}(n \log n)$
%                \item Created a test case generation tool that creates random XML docs from XML Schema
%                \item Automated the extraction and processing of large datasets from legacy systems using SQL and Perl scripts
%            \end{highlights}
%        \end{onecolentry}

    \section{Selected recent publications}
        \begin{samepage}
            \begin{twocolentry}{2024}
                \begin{highlights}
                \item\href{https://doi.org/10.1016/j.ijplas.2024.104098}{Direct application of elasto‑viscoplastic self‑consistent crystal plasticity model to U‑draw bending and springback of dual‑phase high strength steel}
                \end{highlights}
            \end{twocolentry}
            \begin{onecolentry}
                {\it International Journal of Plasticity} 181
            \end{onecolentry}
            \begin{onecolentry}
                B. Jeon, S.-Y. Lee, J. Lee, {\underline{\textit{Y. Jeong}*}}
            \end{onecolentry}
            \vspace{0.10 cm}

            \begin{twocolentry}{2023}
                \begin{highlights}
                \item\href{https://doi.org/10.1016/j.ijmecsci.2022.107796}{A crystal plasticity finite element analysis on the effect of prestrain on springback}
                \end{highlights}
            \end{twocolentry}
            \begin{onecolentry}
                {\it International Journal of Mechanical Sciences} 237
              \end{onecolentry}
            \begin{onecolentry}
                M. Joo, M. -S. Wi, S.-Y. Yoon, S.-Y. Lee, F. Barlat, C. N. Tomé, B. Jeon, {\underline{\textit{Y. Jeong}*}},
             \end{onecolentry}
            \vspace{0.10 cm}

            \begin{twocolentry}{2021}
                \begin{highlights}
                \item\href{https://doi.org/10.1016/j.ijplas.2021.103110}{Finite element analysis using an incremental elasto‑visco‑plastic self‑consistent polycrystal model: FE simulations on Zr and low‑carbon steel subjected to bending, stress‑relaxation, and unloading}
                \end{highlights}
            \end{twocolentry}
            \begin{onecolentry}
                {\it International Journal of Plasticity} 147
            \end{onecolentry}
            \begin{onecolentry}
                \mbox{\underline{\textit{Y. Jeong}*}}, \mbox{B. Jeon}, \mbox{C. N. Tomé}
            \end{onecolentry}
            \vspace{0.10 cm}

            \begin{twocolentry}{2020}
                \begin{highlights}
                \item\href{https://doi.org/10.1016/j.ijplas.2020.102812}{An efficient elasto‑visco‑plastic self‑consistent formulation: Application to steel subjected to loading path changes}
                \end{highlights}
            \end{twocolentry}
            \begin{onecolentry}
                {\it International Journal of Plasticity} 135
            \end{onecolentry}
            \begin{onecolentry}
                {\underline{\textit{Y. Jeong}*}}, C. N. Tomé
            \end{onecolentry}
            \vspace{0.10 cm}
        \end{samepage}

        \section{Goverment-funded projects}
        \begin{twocolentry}{2023 - 2027}
            \textbf{Mid-career Researcher Program (PI), National Research Foundation of Korea}
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{2020 - 2027}
            \textbf{BK21, National Research Foundation of Korea}
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{2018 - 2024}
            \textbf{Engieering Research Center, National Research Foundation of Korea}
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{2020 - 2023}
            \textbf{General Research Program (PI), National Research Foundation of Korea}
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{2018 - 2020}
            \textbf{Virtual engineering platform project, Ministry of Trade, Industry and Energy}
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{twocolentry}{2017 - 2020}
            \textbf{General Research Program (PI), National Research Foundation of Korea}
        \end{twocolentry}

    \section{Skills}
        \begin{onecolentry}
            \item Constitutive modeling: Phenomenological plasticity and crystal plasticity modeling
            \item Programming languages: Fortran, Python (NumPy, SciPy, and matplotlib), Matlab, shell scripts
            \item Languages: Korean, English
            \item Experimental Mechanics: Uniaxial tension, shear, hydraulic bulge test, biaxial tests
            \item Diffraction experiments: X-ray diffraction, Electron back-scattered diffraction and analysis
            \item Computer skills: Linux, Git
        \end{onecolentry}

    \section{GitHub repositories}
        \begin{twocolentry}{\href{https://github.com/youngung/evpsc}{github.com/youngung/evpsc}}\textbf{Elasto-visco-plastic self-consistent model} (private repo)
        \end{twocolentry}
        \vspace{0.10 cm}
        \begin{onecolentry}
            \begin{highlights}
                \item Extended visco-plastic self-consistent model to account for elasticity ($\Delta$EVPSC and $\sigma$EVPSC)
                \item Stand-alone calculation capabilities
                \item Various example shell and python scripts as well as Jupyter notebook files
                \item User material subroutine (UMAT) of Abaqus/standard solver
                \item Written primarily in Fortran with Python and shell scripts
            \end{highlights}
        \end{onecolentry}
        \vspace{0.4 cm}

        \begin{twocolentry}{\href{https://github.com/Jeonbohye/VPSC8-dEVPSC}{github.com/Jeonbohye/VPSC8-dEVPSC}}\textbf{VPSC8 and EVPSC} (private repo)
        \end{twocolentry}
        \begin{onecolentry}
            \begin{highlights}
                \item Combines VPSC8 and EVPSC models
                \item Numerical implementation of VPSC, $\Delta$EVPSC, and $\sigma$EVPSC constitutive model
                \item Three separate build commands for executables of VPSC8, $\Delta$EVPSC, and $\sigma$EVPSC.
            \end{highlights}
        \end{onecolentry}

        \vspace{0.4 cm}
        \begin{twocolentry}{\href{https://github.com/youngung/texture3}{github.com/youngung/texture3}}
            \textbf{In-house Python scripts for texture analysis}
        \end{twocolentry}
        \begin{onecolentry}
            \begin{highlights}
                \item Can plot pole figures of discrete orientations used in VPSC-like crystal plasticity codes
                \item Written entirely in Python with open-sourced libraries including matplotlib, NumPy and SciPy
            \end{highlights}
        \end{onecolentry}
        \vspace{0.2 cm}
        \section{Misc.}
        \begin{twocolentry}{2023-present}
            \begin{highlights}
                \item Board member of Korean Society for Technology of Plasticity
            \end{highlights}
        \end{twocolentry}
        \vspace{0.2 cm}
        \begin{twocolentry}{2019-present}
            \begin{highlights}
                \item Editoiral board member of Korean Journal of Metals and Materials
            \end{highlights}
        \end{twocolentry}
    \vspace{0.2 cm}
    \newpage
    \section{Full list of publications}
    \begin{enumerate}
        \item
        \begin{twocolentry}{submitted}
            \href{}{Role of recovery in the microstructure development and mechanical behavior of a ductile Mg-Zn-Nd-Y-Zr alloy: an analysis using EBSD data and crystal plasticity simulations}
        \end{twocolentry}
        %\begin{onecolentry}
        %    {\it }
        %\end{onecolentry}
        \begin{onecolentry}
            José Victoria-Hernández, {\underline{\textit{Y. Jeong}*}}, Dietmar Letzig
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2025}
            \href{https://doi.org/10.1016/j.jmps.2024.105925}{Modeling deformation, recovery, and recrystallization of tantalum using a higher order elasto-viscoplastic self-consistent model}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Journal of the Mechanics and Physics of Solids}, 194, 105925
        \end{onecolentry}
        \begin{onecolentry}
            I. A. Riyad, B. Clausen, D. J. Savage, {\underline{\textit{Y. Jeong}}}, D. W. Brown, M. Knezevic*
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2024}
            \href{https://doi.org/10.1016/j.jmrt.2024.11.043}{A critical discussion of elasto‑visco‑plastic self‑consistent (EVPSC) models}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Journal of Materials Research Technology}, 33, 7596-7609
        \end{onecolentry}
        \begin{onecolentry}
            B. Jeon, {\underline{\textit{Y. Jeong}*}}, C. N. Tomé
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2024}
            \href{https://doi.org/10.1016/j.ijplas.2024.104098}{Direct application of elasto‑viscoplastic self‑consistent crystal plasticity model to U‑draw bending and springback of dual‑phase high strength steel}
        \end{twocolentry}
        \begin{onecolentry}
            {\it International Journal of Plasticity}, 181, 104098
        \end{onecolentry}
        \begin{onecolentry}
            Bohye Jeon, Shin-Yeong Lee, Jinwoo Lee, {\underline{\textit{Youngung Jeong}*}}
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2024}
            \href{https://doi.org/10.1016/j.jnucmat.2024.155159}{Thermal ratcheting of uranium simulated with a thermo-elasto-visco-plastic self-consistent polycrystal model}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Journal of Nuclear Materials}, 597, 155159
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Youngung Jeong}*}}, Carlos N. Tomé
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2024}
            \href{https://doi.org/10.1016/j.jma.2024.08.009}{Crystal plasticity finite element simulations on extruded Mg-10Gd rod with texture gradient}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Journal of Magnesium and Alloys}, 12 , 3409-3430
        \end{onecolentry}
        \begin{onecolentry}
            Jaeseong Lee, Dirk Steglich, {\underline{\textit{Youngung Jeong}*}}
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2024}
            \href{https://doi.org/10.1016/j.ijplas.2024.104032}{A comprehensive analysis of cermet design and thermal cyclic stability via elasto-viscoplastic crystal plasticity modeling}
        \end{twocolentry}
        \begin{onecolentry}
            {\it International Journal of Plasticity}, 179, 104032
        \end{onecolentry}
        \begin{onecolentry}
            Glenn R Peterson, {\underline{\textit{Youngung Jeong}}}, Carlos N Tomé, Michael D Sangid*
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2024}
            \href{https://doi.org/10.1016/j.msea.2023.145904}{Temperature-dependent behavior of CP-Ti interpreted via self-consistent crystal plasticity simulation}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Materials Science and Engineering: A}, 890, 145904
        \end{onecolentry}
        \begin{onecolentry}
            Bohye Jeon, Min-Su Lee, Tea-Sung Jun, {\underline{\textit{Youngung Jeong}*}}
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2023}
            \href{https://doi.org/10.1016/j.jma.2022.10.017}{Finite element analysis using elasto-visco-plastic self-consistent polycrystal model for E-form Mg sheet subjected to bending}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Journal of Magnesium and Alloys}, 11, 1393-1407
        \end{onecolentry}
        \begin{onecolentry}
            B. Jeon, M. S. Kim, S. H. Choi, {\underline{\textit{Y. Jeong}*}}
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2023}
            \href{https://doi.org/10.1016/j.ijmecsci.2022.107796}{A crystal plasticity finite element analysis on the effect of prestrain on springback}
        \end{twocolentry}
        \begin{onecolentry}
            {\it International Journal of Mechanical Sciences} 237, 107796
        \end{onecolentry}
        \begin{onecolentry}
            M. Joo, M. -S. Wi, S.-Y. Yoon, S.-Y. Lee, F. Barlat, C. N. Tomé, B. Jeon, {\underline{\textit{Y. Jeong}*}}
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2023}
            \href{https://doi.org/10.1016/j.ijplas.2022.103467}{Reconstructing orientation data from the images of IPF maps and ODF sections extracted from the literature: A data-collection method for machine learning}
        \end{twocolentry}
        \begin{onecolentry}
            {\it International Journal of Plasticity}, 159, 103467
        \end{onecolentry}
        \begin{onecolentry}
            L. Kaushik, K.-S. Park, J.-G. Kim, J. Lee, {\underline{\textit{Y. Jeong}}}, S.-H. Choi*
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2022}
            \href{http://dx.doi.org/10.3365/KJMM.2022.60.8.607}{Prediction and validation of stress triaxiality assisted by elasto-visco-plastic polycrystal model}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Korean Journal of Metals and Materials}, 60, 607-618
        \end{onecolentry}
        \begin{onecolentry}
            Jinhwa Park, {\underline{\textit{Youngung Jeong}*}}
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2022}
            \href{https://doi.org/10.1038/s41598-022-07640-3}{In-situ neutron diffraction study of lattice deformation behaviour of commercially pure titanium at cryogenic temperature}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Scientific Reports}, 12, 3719
        \end{onecolentry}
        \begin{onecolentry}
            M.-S. Lee, T. Kawasaki, T. Yamashita, S. Harjo, Y.-T. Hyun, {\underline{\textit{Y. Jeong}}}, T.-S. Jun
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2021}
            \href{https://doi.org/10.1016/j.ijplas.2021.103110}{Finite element analysis using an incremental elasto‑visco‑plastic self‑consistent polycrystal model: FE simulations on Zr and low‑carbon steel subjected to bending, stress‑relaxation, and unloading}
        \end{twocolentry}
        \begin{onecolentry}
            \it{International Journal of Plasticity}, 147, 103110
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Y. Jeong}*}}, \mbox{B. Jeon}, \mbox{C. N. Tomé}
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2020}
            \href{https://doi.org/10.1016/j.ijplas.2020.102812}{An efficient elasto‑visco‑plastic self‑consistent formulation: Application to steel subjected to loading path changes}
        \end{twocolentry}
        \begin{onecolentry}
            \it{International Journal of Plasticity}, 135, 102812
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Y. Jeong}*}}, \mbox{C. N. Tomé}
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2020}
            \href{https://doi.org/10.1016/j.ijmecsci.2020.105680}{Modelling-assisted description of anisotropic edge failure in magnesium sheet alloy under mixed-mode loading}
        \end{twocolentry}
        \begin{onecolentry}
            {\it International Journal of Mechanical Sciences}, 181, 105680
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Y. Jeong}*}}, Dirk Steglich
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2019}
            \href{https://doi.org/10.1088/1361-651X/ab4b66}{Extension of the VPSC model to account for elasto‑visco‑plastic behavior using a
            perturbed viscoplastic approach}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Modelling and Simulation in Materials Science and Engineering}, 27, 085013
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Y. Jeong}*}}, C. N. Tomé
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2019}
            \href{https://doi.org/10.1016/j.msea.2019.138260}{Superior tensile fracture strength of hot isostatically pressed TiC–steel metallic composite fabricated by a novel infiltration}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Materials Science and Engineering: A}, 764, 138260
        \end{onecolentry}
        \begin{onecolentry}
            S. J. Park, {\underline{\textit{Y. Jeong}}}, C. W. Kim, J. H. Lee, S. C. Cho, S. B. Lee, S. K. Lee, D. H. Kim, H. U. Hong*
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2019}
            \href{https://doi.org/10.1007/s12540-019-00292-5}{Enhancement in viscoplastic self-consistent FLD prediction model and its application for austenitic and ferritic stainless steels}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Metals and Materials International}, 25, 1548–1563
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Y. Jeong}*}}, T. Manninen
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2018}
            \href{https://doi.org/10.1016/j.ijplas.2018.07.010}{A crystal plasticity model for describing the anisotropic hardening behavior of steel
            sheets during strain‑path changes }
        \end{twocolentry}
        \begin{onecolentry}
            {\it International Journal of Plasticity},
        \end{onecolentry}
        \begin{onecolentry}
            H. Kim, F. Barlat, Y. Lee, S. Zaman, C. S. Lee, {\underline{\textit{Y. Jeong}*}}
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2017}
            \href{http://dx.doi.org/10.1016/j.ijplas.2016.07.015}{A comparative study between micro‑ and macro‑mechanical constitutive models developed for complex loading scenarios }
        \end{twocolentry}
        \begin{onecolentry}
            {\it International Journal of Plasticity}, 93, 212‑228
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Y. Jeong}*}}, F. Barlat, C. N. Tomé, W. Wen
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2016}
            \href{https://doi.org/10.1107/S1600576716013662}{Uncertainty in flow stress measurements using X‑ray diffraction for sheet metals subjected to large plastic deformations}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Journal of Applied Crystallography}, 49, 1991‑2004
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Y. Jeong}*}}, T. Gnäupel‑Herold, M. Iadicola, A. Creuziger
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2016}
            \href{http://dx.doi.org/10.1016/j.ijmecsci.2016.08.013}{Texture‑based forming limit prediction for Mg sheet alloys ZE10 and AZ31 }
        \end{twocolentry}
        \begin{onecolentry}
            {\it International Journal of Mechanical Sciences}, 117, 102‑114
        \end{onecolentry}
        \begin{onecolentry}
            D. Steglich, {\underline{\textit{Y. Jeong}*}}
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2016}
            \href{http://dx.doi.org/10.1088/0965-0393/24/5/055005}{Forming limit prediction using a self‑consistent crystal plasticity framework: a case study for BCC fiber textures}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Modelling and Simulation in Materials Science and Engineering}, 24, 055002
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Y. Jeong}*}}, M.‑S. Pham, M. Iaidocola, A. Creuziger, T. Foecke
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2016}
            \href{http://dx.doi.org/10.1016/j.actamat.2016.04.013}{Multiaxial constitutive behavior of an interstitial‑free steel: measurements through X‑ray and digital image correlation }
        \end{twocolentry}
        \begin{onecolentry}
            {\it Acta Materialia}, 112, 84‑93
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Y. Jeong}*}},  M. Iaidocola, T. Gnäupel‑Herold, A. Creuziger
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2016}
            \href{http://dx.doi.org/10.1016/j.msea.2015.09.108}{Effect of martensitic phase transformation on the behavior of 304 austenitic stainless steel under tension}
        \end{twocolentry}
        \begin{onecolentry}
            {\it Materials Science and Engineering: A}, 649, 174‑183
        \end{onecolentry}
        \begin{onecolentry}
            H. Wang*, {\underline{\textit{Y. Jeong}*}}, B. Calusen, Y. LiU, R. J. McCabe, F. Barlat, C. N. Tomé
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2015}
            \href{http://dx.doi.org/10.1016/j.ijplas.2014.06.009}{Evaluation of biaxial flow stress based on Elasto‑Viscoplastic Self‑Consistent analysis of X‑ray Diffraction Measurements}
        \end{twocolentry}
        \begin{onecolentry}
            {\it International Journal of Plasticity}, 66, 103‑118
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Y. Jeong}}}, T. Gnaupel-Herold, F. Barlat, M. Iadicola, A. Creuziger, M.‑G. Lee*
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2012}
            \href{http://dx.doi.org/10.1088/0965-0393/20/2/024009}{Application of crystal plasticity to an austenitic stainless steel }
        \end{twocolentry}
        \begin{onecolentry}
            {\it Modelling and Simulation in Materials Science and Engineering}, 20, 024009
        \end{onecolentry}
        \begin{onecolentry}
            {\underline{\textit{Y. Jeong}*}}, F. Barlat, M.-G. Lee
        \end{onecolentry}
        \vspace{0.10 cm}

        \item
        \begin{twocolentry}{2012}
            \href{http://dx.doi.org/10.1016/j.ijsolstr.2012.06.017}{Biaxial Deformation Behavior of AZ31 Magnesium Alloy: Crystal‑Plasticity‑Based Prediction and Experimental Validation}
        \end{twocolentry}
        \begin{onecolentry}
            {\it International Journal of Solids and Structure}, 49, 3551‑3561
        \end{onecolentry}
        \begin{onecolentry}
            D. Steglich*, {\underline{\textit{Y. Jeong}}}, M. O. Andar, T. Kuwabara
        \end{onecolentry}
        \vspace{0.10 cm}

    \end{enumerate}

\end{document}